Statistical Theory\\ Chess Dataset Analysis

### 2024-09-29 Article/Article.tex
Updated Article

\subsection{Rating Distribution}
The distribution of player ratings in the dataset initially resembles a normal distribution with a peak around the mean. However, this peak diminishes when games involving unrated players are excluded, bringing the distribution closer to normality. This is illustrated in Figure~\ref{fig:transformation_plot}. Additionally, the QQ plot (Figure~\ref{fig:rating_qq}) supports the similarity between the player rating distribution and a normal distribution, albeit with deviations at the extremes.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\linewidth]{ratingqq.png}
    \caption{Quantile - Quantile plot of player rating in relation to fitted normal distribution}
    \label{fig:rating_qq}
\end{figure}

@@

@@

\subsection{Relations Between Features}
Analyzing the dataset’s convolution matrix reveals linear relationships between certain features. For example, there is a strong correlation between the ratings of white and black players in a game, which aligns with the expectation that players of similar ratings tend to compete against each other. However, no significant linear relationship was observed between the winner and the ratings of either player.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\linewidth]{ratingecdf.png}
    \caption{Gamma-Poisson distribution fitted to the 'Turns' data}
    \label{fig:ecdf_plot}
\end{figure}

\subsection{Summary of Results}
The analysis demonstrates a clear statistical relationship between player rating and game length. The Mann-Whitney U test confirmed that higher-rated players play longer games. This finding supports the hypothesis that player rating is a factor in game duration, with the distribution of ratings further informing our understanding of game dynamics.

>>>>

\subsection{Rating Distribution}
The distribution of player ratings in the dataset initially resembles a normal distribution with a peak around the mean. However, this peak diminishes when games involving unrated players are excluded, bringing the distribution closer to normality. This is illustrated in Figure~\ref{fig:transformation_plot}. Additionally, the QQ plot (Figure~\ref{fig:rating_qq}) supports the similarity between the player rating distribution and a normal distribution, albeit with deviations at the tails, which are exacerbated by the QQ plot's sensitivity to deviations at the tail.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\linewidth]{ratingqq.png}
    \caption{Quantile - Quantile plot of white player rating in relation to normal distribution}
    \label{fig:rating_qq}
\end{figure}

@@

\subsection{Relations Between Features}
Analyzing the dataset’s correlation matrix reveals linear relationships between certain features. For example, there is a weak positive correlation between the players' average rating and game length. There are also weak negative correlations between the number of turns and both the absolute difference in player ratings and the 'winner' variable, where 'winner' is numerized as 0 for draws and 1 for games with a winner. From these correlations, we can tentatively conclude that:
\begin{itemize}
    \item Games involving more skilled players tend to last longer.

    \item Games with a larger skill gap between players may end faster.

    \item Draws tend to draw out longer then desicive games.
\end{itemize}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\linewidth]{corr_mat.png}
    \caption{Heat map of the correlation matrix}
    \label{fig:corr_mat}
\end{figure}

@@
